---
title: "Pareto"
output: html_vignette
vignette: >
  %\VignetteIndexEntry{Pareto}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  # eval = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.width=8, 
  fig.height=5, 
  fig.align = "center"
)
```


```{r setup}
library(Morphoscape)
library(rPref)

library(ape)
library(phytools)
# Load turtle phylogeny and new turtle dataframe names
tTree <- ape::read.tree("C:/Users/Rob/Documents/Code Projects/Morphoscape/Turtle_Tree.nwk")
new_names <- readLines("C:/Users/Rob/Documents/Code Projects/Morphoscape/Turtle_List.txt")

data(warps)
data(turtles)
# Update rownames based on updated taxonomy
rownames(turtles) <- new_names
```

The following provides a guide on the workflow of performing adaptive landscape analysis using Pareto optimality.

If you haven't already, please read the accompanying "Morphoscape" vignette for an introduction to adaptive landscapes more generally aas well as some core functions of the package!

# Considering Optimality
## Functional Optimality

Write something here about functional adaptive landscapes and quickly go through the Morphoscape workflow...

```{r}
data("turtles")
data("warps")

str(turtles)
str(warps)

# Make performance surfaces
warps_fnc <- as_fnc_df(warps, func.names = c("hydro", "curve", "mech", "fea"))
kr_surf <- krige_surf(warps_fnc, hull = FALSE, new_data = turtles)
# Make functional adaptive landscapes
weights <- generate_weights(n = 20, data = kr_surf)
all_landscapes <- calc_all_lscps(kr_surf, grid_weights = weights)
wprime_by_Group <- calcWprimeBy(all_landscapes, by = ~Ecology,method = "quantile")
```

Show off new functions for extracting data for custom plotting (e.g., on phylogenies)

```{r}
# Get heights on each landscape for individual species
Landscape_Values <- sp_vals_from_grid(wprime_by_Group,turtles)
# Plot landscape heights for different groups
# Here using 'marine' as an example
ggplot(Landscape_Values,aes(x=Ecology, y=`M`))+
  geom_boxplot()
# Plot landscape heights on phylogeny
# Here using 'marine' as an example
trait_2_plot <- Landscape_Values$M
names(trait_2_plot) <- rownames(Landscape_Values)
phytools::contMap(tTree,trait_2_plot)
```

## Pareto Optimality

An alternative way of thinking about optimality comes from Pareto optimality. A structure might have to compromise between different functions due to inherent trade-offs. See the differences in the performance surfaces for two traits from the turtle dataset.
```{r}
# Make performance surfaces for individual traits
plot_ggsurf(kr_surf,'hydro')
plot_ggsurf(kr_surf,'curve')
```

While trade-offs inevitably exist, some morphologies handle them better than others. For example, a trade-off exists between curvature and hydrodynamic performance, and you cannot simultaeneously maximize both traits. 

However, you can still maximize one trait *for a given value* of the second trait; morphologies that meet this criterion are referred to as 'Pareto optimal', and the paathway they make through morphospace is referred to as the 'Pareto front'.
```{r}
# Compute Pareto landscape for two traits
KrSurfParOpt <- calc_lscp_Pareto(kr_surf,'hydro','curve')
# Plot trade-offs between the traits
plot_ggtrade(KrSurfParOpt)
# Plot the Pareto landscape
plot_ggPareto(KrSurfParOpt)

# Get Pareto optimality values for individual species
Ri_Values <- sp_vals_from_grid(KrSurfParOpt,turtles)

# Plot species Pareto optimality values for different groups
ggplot(Ri_Values,aes(x=Ecology, y=`hydro-curve`))+
  geom_boxplot()
# Plot species Pareto optimality values on a phylogeny
trait_2_plot <- Ri_Values$`hydro-curve`
names(trait_2_plot) <- rownames(Ri_Values)
phytools::contMap(tTree,trait_2_plot)
```

Pareto optimality is generally defined in terms of trade-offs between traits, hence why the example above used performance surfaces as the input. 

However, we can extend the same methods to a more abstract case, and examine trade-offs between performance on different adaptive landscapes, as shown below.

```{r}
#########
# Plot individual group landscapes
plot_ggsurf(wprime_by_Group,'M')
plot_ggsurf(wprime_by_Group,'S')

# Calculate Pareto landscape based on functional landscapes
Lscp_Pareto_MS <- calc_lscp_Pareto(wprime_by_Group,'M','S')
summary(Lscp_Pareto_MS)
# Plot trade-offs between landscape heights
plot_ggtrade(wprime_by_Group,'M','S')
# Plot Pareto landscape between landscape heights
plot_ggPareto(wprime_by_Group,'M','S')
```

We can even do all of that with just one function!

```{r}
plot_lscp_Pareto(wprime_by_Group,'M','S')
```



## References

Bookstein, F. L. (2019). Pathologies of between-groups principal components analysis in geometric morphometrics. Evolutionary Biology, 46(4), 271-302.

Cardini, A., O’Higgins, P., & Rohlf, F. J. (2019). Seeing distinct groups where there are none: spurious patterns from between-group PCA. Evolutionary Biology, 46(4), 303-316.

Cardini, A., & Polly, P. D. (2020). Cross-validated between group PCA scatterplots: A solution to spurious group separation?. Evolutionary Biology, 47(1), 85-95.

Dickson, B. V., & Pierce, S. E. (2019). Functional performance of turtle humerus shape across an ecological adaptive landscape. Evolution, 73(6), 1265-1277.

Smith, S. M., Stayton, C. T., & Angielczyk, K. D. (2021). How many trees to see the forest? Assessing the effects of morphospace coverage and sample size in performance surface analysis. Methods in Ecology and Evolution, 12(8), 1411-1424.
